National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs

### 2021-03-11 Supplement_And_Analysis.Rmd
updated manuscript

Informal analysis indicates that unsolicited opt-ins were often the partners of existing participants. The sample retained 4,053 participants from the initial Time 1 (2009) NZAVS of 6,518 participants (a retention rate of 62.2% over three years). The sample retained 5,762 participants from the full Time 3 (2011) sample (a retention rate of 83.7% from the previous year). Participants were posted a copy of the questionnaire, with a second postal follow-up two months later. Participants who provided an email address were also emailed and invited to complete an online version if they preferred. 
Non-respondents were emailed a follow-up reminder email approximately two months later. Three attempts were then made using each provided phone number (typically home and cell phone) to contact non-respondents to encourage participation. These attempts were made on separate days, approximately one week apart. When possible, a phone message was left for each phone number after the third attempt. Participants were also posted a pamphlet outlining recent findings from the study mid-way through the year. Finally, participants were posted a Season’s Greetings from the NZAVS research team, and informed that they had been automatically entered into a bonus seasonal grocery voucher prize draw for a total pool of $NZ 1,000). Participants were informed that the draw would happen automatically and winners contacted. The Season’s Greetings card also asked participants to contact us (online, email or phone) to let us know if any of their contact details had changed before the prize draw was conducted. These additional materials are presented by Huang, Greaves, and Sibley (2014) in an online NZAVS technical report.

@@

Time 5 indicated that the longer length of this questionnaire may have contributed to the low response rate in this case. 
The third sample frame consisted of 3,000 people randomly selected from the 2012 Electoral Roll who lived in the Christchurch region. A total of 332 participants responded to this booster sample (adjusted response rate = 11.24%). The Christchurch region was oversampled because it has experienced significant hardship and change due to the Christchurch earthquakes of 2010 and 2011 with many people moving out of the region (Statistics New Zealand, 2013) and problems with mail delivery with some city zones being placed under restricted entry due to safety concerns and considerable infrastructure destroyed). 
The fourth sample frame consisted of 9,000 respondents selected from meshblock area units across the country that were moderate-to-high in deprivation according to the index developed by Salmond, Crampton and Atkinson (2007). Regions with levels of deprivation were selected using scores on the decile-ranked NZ Deprivation index from 6-10, with 10 being the most deprived). This sample frame used scaled weighting so that people in increasingly deprived regions were increasingly more likely to be selected (with random sampling of people within regions that had a given level of deprivation). The scaling factor was as follows: ni = nbase * weighti, where nbase = 600, and weighti, ranged from 1 to 5 and increased by 1 for each one-unit increased in deprivation score. Thus, 600 people were randomly selected from regions with a deprivation score of 6, 1,200 people were randomly selected from regions with a deprivation score of 7, and so on. This sampling strategy was designed to increase the representativeness of the sample across regions with different levels of deprivation, as the NZAVS showed increased attrition in increasingly more deprived regions over the first three years of the study. A total of 767 participants responded to this booster sample (adjusted response rate = 8.65%). The fifth sample frame consisted of 9,000 people randomly selected from those who indicated on the 2012 Electoral Roll that they were of Māori ethnicity (ethnic affiliation as Māori is listed on the roll, but other ethnic affiliations are not). A total of 689 participants responded to this booster sample (adjusted response rate = 7.78%). The questionnaire administered to the Māori booster sample included questions specifically designed for Māori.

@@

The Time 11 (2019) NZAVS contained responses from 42,684 participants (36,522 retained from one or more previous wave. The sample retained 2,506 participants from the Time 1 (2009) sample (a retention rate of 38.4%). The sample retained 34,782 participants from Time 10 (2018; a retention rate of 72.5% from the previous year). Participants who provided an email address were first emailed and invited to complete an online version if they preferred. Participants who did not complete the online version (or did not provide an email) were then posted a copy of the questionnaire, with a second postal follow-up two months later. We staggered the time of contact, so that participants who had completed the previous wave were contacted approximately one year after they last completed the questionnaire. A second reminder email was sent approximately four months following the first email attempt. We offered a prize draw for participation (five draws each for $1000 grocery vouchers, $5000 total prize pool). Participants were also emailed an eight-page newsletter about the study. As in past years, three attempts were made to phone non-respondents using each available cell and landline number. However, due to the university closure during COVID-19 lockdowns, phoning attempts were made for only 54% of the phoning pool (11,687 from a total of 21,636 non-respondents who provided at least one phone number).

@@

The Facebook advertisement read as follows: “Participate in the New Zealand Attitudes and Values Study. Complete the 2020 Questionnaire online” with the body of text: “If you are part of the NZAVS, but have not heard from us in the last year, then please consider completing the 2020 questionnaire online. The study is more important than ever as we aim to understand the impact of COVID-19 on mental health, wellbeing and resilience in our communities. We wish you all the best at this time and hope you keep well and stay safe.” This paid promotion reached 883,969 people, with 37,850 link clicks (i.e., clicking the link for the Qualtrics survey) according to Facebook. A total of 6106 people continued complete the questionnaire and provide full contact details, and were thus included in the dataset (4734 were new participants opting in to the study, and 1372 were previously ‘lost’ participants).

@@

This supplement offers a step-by-step analysis for the "Longitudinal
Evidence for Growth in Subjective Well-being from Spiritual Beliefs."
The functions and other materials for this analysis can be found at the
following GitHub link:
<https://github.com/jacopastorius/nzavs_spiritbeliefwell-being>

@@

# pa<-d%>%
#  dplyr::filter(YearMeasured==1)%>%
#     dplyr::group_by(Id) %>%
#     dplyr::filter(n() !=0)%>%
#     dplyr::ungroup(Id)
# length(unique(pa$Id))
# table(pa$Wave)

# d2<-data_clean_spirit_wellbeing(pa,0) # d = dataframe, 3 = number of waves
# table(d2$Wave)

# function to clean data, specify number of waves as 3 
d_3<-data_clean_spirit_wellbeing(d,9) # d = dataframe, 3 = number of waves
# function to count uinque ides
show_unique_id(d_3$Id) 
# function to count waves for participants
count_waves_participants(d_3)
# table of beliefs 
table(d_3$Beliefs)
# years Note that we use a continuous response measure participants. 
range(d_3$years)
# html talbe
#demographic_table(d_3)

d_3 <- cbind(
  d_3,
  demean(d_3, select = c("Age"), group = "Id"
))

@@

@@

@@

<!-- ```{r  badlatextable} badd tables -->

<!-- library("arsenal") -->

<!-- summary(arsenal::tableby(Wave ~ Age + -->

<!--                    NZdep + -->

<!--                    Education + -->

<!--                    Employed + -->

<!--                    Ethnic_Categories + -->

<!--                    Male  + -->

<!--                    Has_Partner  + -->

<!--                    Political_Conservativism + -->

<!--                    Urban + -->

<!--                    LIFESAT + PWI + includeNA(Beliefs), data = d_3,digits = 2, total=FALSE, format.p = FALSE, digits.p = 0))%>% -->
<!--     kable( booktabs = T, "latex") -->
<!-- ``` -->

@@

<!-- ```{r, include=FALSE} -->

<!-- beliefs_table(d_3) -->

<!-- We wrote the publication in $\LaTeX$, which requires converting the tables to .tex files; this conversion used a custom function we wrote in the `funs.R` file.  -->

<!-- ```{r latex_data_table, include=FALSE} -->

<!-- # you might need to install the dev verion.//remotes::install_github("tysonstanley/furniture") -->

<!-- demographic_table_latex() # note the furniture package conflicts with the table1 package. -->

<!-- d_3%>% -->

<!--   group_by(Wave)%>% -->

<!-- furniture::table1( -->

<!--                   Age,  -->

<!--                   NZdep, -->

<!--                   Education, -->

<!--                   Employed, -->

<!--                   Ethnic_Categories, -->

<!--                   Male,Has_Partner,  -->

<!--                   Political_Conservativism, -->

<!--                   Urban,  -->

<!--                   Beliefs,  -->

<!--                   LIFESAT, -->

<!--                   PWI, -->

<!--                   overall=F) -->

<!-- dat3<-d_3%>%dplyr::select( Age, NZdep, -->

<!--                   Education, -->

<!--                   Employed, -->

<!--                   Ethnic_Categories, -->

<!--                   Male,Has_Partner, -->

<!--                   Political_Conservativism, -->

<!--                   Urban, -->

<!--                   Beliefs, -->

<!--                   LIFESAT, -->

<!--                   PWI,Wave) -->

<!-- stargazer::stargazer(dat3) -->

<!-- library(modelsummary) -->

<!--   modelsummary::datasummary_df(ALL(Age) ~ Mean + SD + Histogram, data=dat3) -->

<!-- ## How many waves to select? We are using 10 years -->

<!-- We selected a minimum of three waves because we want to recover a -->

<!-- longitudinal sample in which a slope for the effect of the years -->

<!-- $\times$ beliefs effect might be recovered. However, we needed to -->

<!-- consider whether our selection of three waves was sensitive to that -->

<!-- degree of researcher freedom. So we also prepared 4-8 years of data -->

<!-- frames for separate analyses. -->

<!-- ```{r, include = TRUE} -->

<!-- # create datasets to test robustness to years -->

<!-- # d_4<-data_clean_spirit_wellbeing(d,4) # 4 years -->

<!-- # d_5<-data_clean_spirit_wellbeing(d,5) # 5 years -->

<!-- # d_6<-data_clean_spirit_wellbeing(d,6) # 6 years -->

<!-- # d_7<-data_clean_spirit_wellbeing(d,7) # 7 years -->

<!-- # d_8<-data_clean_spirit_wellbeing(d,8) # 7 years -->

<!-- # # function to count uinque number of ID's -->

<!-- # show_unique_id(d_3$Id)  -->

<!-- # show_unique_id(d_4$Id) -->

<!-- # show_unique_id(d_5$Id)  -->

<!-- # show_unique_id(d_6$Id)  -->

<!-- # show_unique_id(d_7$Id)  -->

<!-- # show_unique_id(d_8$Id)  -->

<!-- # # function to count waves for participants -->

<!-- # count_waves_participants(d_3) -->

<!-- # count_waves_participants(d_4) -->

<!-- # count_waves_participants(d_5) -->

<!-- # count_waves_participants(d_6) -->

<!-- # count_waves_participants(d_7) -->

<!-- # count_waves_participants(d_8) -->

<!-- The analyses for these waves is described below. These analyses reveal -->

<!-- that the inference of the paper is not sensitive to the threshold number -->

<!-- of waves for participant selection. Which is to say, we recover the same -->

<!-- inference with an -wave selection criterion. -->

@@

# theoretically plausible relationships
tidy_ggdag <- dagify(
  WB ~ belief + age_within + age_between + partner + nzdep + urban + male + pols + empl, # what might affect wellbeing?
  WB ~~ partner, # bi-causal path
  belief ~ age_within  + age_between  + male + ethn, # predictors of beliefs
  partner ~ nzdep + age_within  + age_between + belief,  # predictors of having a partner
  nzdep ~ empl,  # predictor of deprivation
  pols ~ age_within  + age_between  + empl + ethn, # predictor of political orientation
  empl ~ nzdep + edu + ethn + age_within  + age_between, # predictor of employment
  exposure =  "belief",
  outcome =   "WB")%>%
  tidy_dagitty() + theme_dag_gray()

#graph
ggdag_adjustment_set(tidy_ggdag, node_size = 14) + 
  theme(legend.position = "bottom") 
```

@@

@@

@@

Next we plot the graphs. First we extract the predictions

```{r Bayesian_pwi_predictedSAT_DAG_GET, cache=TRUE}
get_predictions(mod.1_bDAGR)
```

@@

<!-- We might want to look at within/between indivual change (i.e. do people who change in degree of spirit beliefs manifest different well-being.  I ran this as below, however, what we really want is a markov model, where we compare state changes between the different types of belief. So this model is not useful. For more on demeaning see:  -->

>>>>

Informal analysis indicates that unsolicited opt-ins were often the partners of existing participants. The sample retained 4,053 participants from the initial Time 1 (2009) NZAVS of 6,518 participants (a retention rate of 62.2% over three years). The sample retained 5,762 participants from the full Time 3 (2011) sample (a retention rate of 83.7% from the previous year). Participants were posted a copy of the questionnaire, with a second postal follow-up two months later. Participants who provided an email address were also emailed and invited to complete an online version if they preferred.

Non-respondents were emailed a follow-up reminder email approximately two months later. Three attempts were then made using each provided phone number (typically home and cell phone) to contact non-respondents to encourage participation. These attempts were made on separate days, approximately one week apart. When possible, a phone message was left for each phone number after the third attempt. Participants were also posted a pamphlet outlining recent findings from the study mid-way through the year. Finally, participants were posted a Season’s Greetings from the NZAVS research team, and informed that they had been automatically entered into a bonus seasonal grocery voucher prize draw for a total pool of $NZ 1,000). Participants were informed that the draw would happen automatically and winners contacted. The Season’s Greetings card also asked participants to contact us (online, email or phone) to let us know if any of their contact details had changed before the prize draw was conducted. These additional materials are presented by Huang, Greaves, and Sibley (2014) in an online NZAVS technical report.

@@

Time 5 indicated that the longer length of this questionnaire may have contributed to the low response rate in this case. 
The third sample frame consisted of 3,000 people randomly selected from the 2012 Electoral Roll who lived in the Christchurch region. A total of 332 participants responded to this booster sample (adjusted response rate = 11.24%). The Christchurch region was oversampled because it has experienced significant hardship and change due to the Christchurch earthquakes of 2010 and 2011 with many people moving out of the region (Statistics New Zealand, 2013) and problems with mail delivery with some city zones being placed under restricted entry due to safety concerns and considerable infrastructure destroyed). 
The fourth sample frame consisted of 9,000 respondents selected from meshblock area units across the country that were moderate-to-high in deprivation according to the index developed by Salmond, Crampton and Atkinson (2007). Regions with levels of deprivation were selected using scores on the decile-ranked NZ Deprivation index from 6-10, with 10 being the most deprived). This sample frame used scaled weighting so that people in increasingly deprived regions were increasingly more likely to be selected (with random sampling of people within regions that had a given level of deprivation). The scaling factor was as follows: n_i = n_base * weigh_i, where n_base = 600, and weight_i, ranged from 1 to 5 and increased by 1 for each one-unit increased in deprivation score. Thus, 600 people were randomly selected from regions with a deprivation score of 6, 1,200 people were randomly selected from regions with a deprivation score of 7, and so on. This sampling strategy was designed to increase the representativeness of the sample across regions with different levels of deprivation, as the NZAVS showed increased attrition in increasingly more deprived regions over the first three years of the study. A total of 767 participants responded to this booster sample (adjusted response rate = 8.65%). The fifth sample frame consisted of 9,000 people randomly selected from those who indicated on the 2012 Electoral Roll that they were of Māori ethnicity (ethnic affiliation as Māori is listed on the roll, but other ethnic affiliations are not). A total of 689 participants responded to this booster sample (adjusted response rate = 7.78%). The questionnaire administered to the Māori booster sample included questions specifically designed for Māori.

@@

The Time 11 (2019) NZAVS contained responses from 42,684 participants (36,522 retained from one or more previous wave. The sample retained 2,506 participants from the Time 1 (2009) sample (a retention rate of 38.4%). The sample retained 34,782 participants from Time 10 (2018; a retention rate of 72.5% from the previous year). Participants who provided an email address were first emailed and invited to complete an online version if they preferred. Participants who did not complete the online version (or did not provide an email) were then posted a copy of the questionnaire, with a second postal follow-up two months later. We staggered the time of contact, so that participants who had completed the previous wave were contacted approximately one year after they last completed the questionnaire. A second reminder email was sent approximately four months following the first email attempt. We offered a prize draw for participation (five draws each for \$1000 grocery vouchers, \$5000 total prize pool). Participants were also emailed an eight-page newsletter about the study. As in past years, three attempts were made to phone non-respondents using each available cell and landline number. However, due to the university closure during COVID-19 lockdowns, phoning attempts were made for only 54% of the phoning pool (11,687 from a total of 21,636 non-respondents who provided at least one phone number).

@@

The Facebook advertisement read as follows: “Participate in the New Zealand Attitudes and Values Study. Complete the 2020 Questionnaire online” with the body of text: "If you are part of the NZAVS, but have not heard from us in the last year, then please consider completing the 2020 questionnaire online. The study is more important than ever as we aim to understand the impact of COVID-19 on mental health, wellbeing and resilience in our communities. We wish you all the best at this time and hope you keep well and stay safe." This paid promotion reached 883,969 people, with 37,850 link clicks (i.e., clicking the link for the Qualtrics survey) according to Facebook. A total of 6106 people continued complete the questionnaire and provide full contact details, and were thus included in the dataset (4734 were new participants opting in to the study, and 1372 were previously ‘lost’ participants).

@@

This supplement offers a step-by-step analysis for the manuscript: "Longitudinal
Evidence for Growth in Subjective Well-being from Spiritual Beliefs."

The functions and figures for this analysis can be found at the
following GitHub link:
<https://github.com/go-bayes/nzavs_spiritbeliefwell-being>

@@

# function to clean data, specify number of waves as 3 
d_3<-data_clean_spirit_wellbeing(d,9) # d = dataframe, 9 = number of waves
d_3a<-data_clean_spirit_wellbeing(d,3) # d = dataframe, 3 = number of waves

# function to count uinque ides
show_unique_id(d_3$Id) 
# function to count waves for participants
count_waves_participants(d_3)
# table of beliefs 
table(d_3$Beliefs)
# years Note that we use a continuous response measure participants. 
range(d_3$years)
# html talbe
#demographic_table(d_3)

## For data with min 9 x waves
d_3 <- cbind(
  d_3,
  demean(d_3, select = c("Age"), group = "Id"
))

@@

# data with 3 x waves.
d_3a <- cbind(
  d_3a,
  demean(d_3a, select = c("Age"), group = "Id"
))

d_3a<-d_3a%>%
  dplyr::mutate(Age_betweenC = scale(Age_between, scale = FALSE))
```

@@

# Unique participants

```{r}
length(unique(d_3$Id)) #3257
```

@@

@@

@@

# theoretically plausible relationships
tidy_ggdag <- ggdag::dagify(
  WB ~ belief + age_within + age_between + partner + nzdep + urban + male + pols + empl,
  WB ~~ partner,
  belief ~ age_within  + age_between  + male + ethn,
  partner ~ nzdep + age_within  + age_between + belief,
  nzdep ~ empl,
  pols ~ age_within  + age_between  + empl + ethn,
  empl ~ nzdep + edu + ethn + age_within  + age_between,
  exposure =  "belief",
  outcome =   "WB"
) %>%
  tidy_dagitty()

ggdag_adjustment_set(tidy_ggdag, node_size = 14) + 
  theme(legend.position = "bottom")  + theme_dag_gray()
```

@@

## Models

```{r}
# REDO
m1<-brm(forumula_bayesPWI_DAG, data= d_3, file ="mod.1_bDAGR.rds", cores = 8) # note these are very large files
summary(m1)
```

```{r}
conditions <- data.frame(Age_within = c(-5:5))
p0 <- brms::conditional_effects(
  m1,
  c("Age_within:Beliefs" ),
  spaghetti = TRUE,
  categorical = FALSE,
  nsamples = 200,
  prob = .95,
  method = 'fitted',
  conditions = conditions
)

plot(p0, plot = FALSE)[[1]] +   facet_wrap("Beliefs",nrow = 1) + xlab("Years of Aging (centered)") + ylab ("Personal Wellbeing 0-10")  + theme_blank() + labs(title = "Predicted effects of beliefs states for personal wellbeing")+ theme(legend.position="top") + scale_color_viridis_d(alpha = .2)
# pwi

```{r lsbayesmodelLS_DAG2}
# LIFE SAT
m2 <-brm (forumula_bayesLIFESAT_DAG, data= d_3,cores = 8, file ="mod.2_bDAGR.rds")
summary(m2)
###
# note these are very large files
```

@@

```{r}

@@

Next we plot the graphs. First we extract the predictions:

```{r Bayesian_pwi_predictedSAT_DAG, cache=TRUE}
get_predictions(m1)
```

```{r Bayesian_lifesat_predictedSAT_DAG_GET2, cache=TRUE}
get_predictions(mod.2_bDAGR)
```
```{r Bayesian_pwi_predictedSAT_DAG_GET, cache=TRUE}
get_predictions(mod.1_bDAGR)
```

@@

<!-- We might want to look at within/between individual change (i.e. do people who change in degree of spirit beliefs manifest different well-being.  I ran this as below, however, what we really want is a markov model, where we compare state changes between the different types of belief. So this model is not useful. For more on demeaning see:  -->
